---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---
## Loading and preprocessing the data
```{r}
dat<-read.csv('activity.csv')
```
## What is mean total number of steps taken per day?
```{r plot,fig.height=5}
stepsPerDay<-sapply(split(dat$steps,dat$date),sum)
hist(stepsPerDay)
```
```{r}
meanSteps<-mean(stepsPerDay[!is.na(stepsPerDay)])
medianSteps<-median(stepsPerDay[!is.na(stepsPerDay)])
```
The mean steps per day is `r meanSteps` and the median steps per day is `r medianSteps`.  

## What is the average daily activity pattern?
```{r plotmean,fig.height=4}
dailyMeanActivity<-tapply(dat$step,dat$interval,mean,na.rm=TRUE)
intervals=levels(factor(dat$interval))
plot(intervals,dailyMeanActivity,type="l",main='Daily Mean Activity')
```
```{r}
maxInd<-match(max(dailyMeanActivity),dailyMeanActivity)
maxInterval<-intervals[maxInd]
```
The 5 minute interval containing the maximum number of steps is at `r maxInterval`  
## Inputing missing values
```{r histplot,fig.height=5}
numNAs<-length(dat[is.na(dat$steps)==TRUE,'steps'])
datFilled<-dat
dailyMeanVec<-rep(dailyMeanActivity,length=length(dat$steps))
for(i in 1:length(dat$steps)){
        if(is.na(datFilled[i,'steps'])==TRUE){
                datFilled[i,'steps']<-dailyMeanVec[i]
        }
}

stepsPerDayFilled<-sapply(split(datFilled$steps,datFilled$date),sum)
hist(stepsPerDayFilled)
```
```{r}
meanStepsFill<-mean(stepsPerDayFilled[!is.na(stepsPerDayFilled)])
medianStepsFill<-median(stepsPerDayFilled[!is.na(stepsPerDayFilled)])
meanStepsDiff<-abs(meanStepsFill-meanSteps)
medianStepsDiff<-abs(medianStepsFill-medianSteps)
```
Mean steps taken per day is `r meanStepsFill`. Median steps taken per day is `r medianStepsFill`.  
These two have a difference of `r meanStepsDiff` and `r medianStepsDiff` from the nonfilled mean and median, respectively. These two differences are negligible and thus the values do not differ.  

## Are there differences in activity patterns between weekdays and weekends?
```{r plotcomparison,fig.height=8}
week<-sapply(as.Date(dat$date),weekdays)
isWeekend<-rep(c('Weekday','Weekend'),length=length(week))
for(i in 1:length(week)){
        if(week[i]=='Monday'| week[i]=='Tuesday'| week[i]=='Wednesday'| week[i]=='Thursday'| week[i]=='Friday'){
                isWeekend[i]<-'Weekday'
                } else{
                        isWeekend[i]<-'Weekend'
                }
}
dat<-data.frame(dat,isWeekend)
weekdayMeanActivity<-tapply(dat[dat['isWeekend']=='Weekday','steps'],dat[dat['isWeekend']=='Weekday','interval'],mean,na.rm=TRUE)
weekendMeanActivity<-tapply(dat[dat['isWeekend']=='Weekend','steps'],dat[dat['isWeekend']=='Weekend','interval'],mean,na.rm=TRUE)

par(mfrow=c(2,1))
plot(intervals,weekdayMeanActivity,type='l',main='Weekday')
plot(intervals,weekendMeanActivity,type='l',main='Weekend')
```
